Fusing Filters with Integer Linear Programming

### 2014-05-16 2014betterfusionforfilters/section/04-ILP.tex
erik suggestion on abstract
rename split to possible - there is a possibility they may be fused

\begin{figure}
\begin{tabbing}
MMMMM       \= M  \= \kill
$nodes$     \> @:@ \> $program \to V$              \\
$edges$     \> @:@ \> $program \to E$              \\
$edge$      \> @:@ \> $\{bind\} \times bind \to E$ 
\\[1ex]
$nodes(bs)$ \> $= \{(name(b), \iiter_{\Gamma,C}(b)) | b \in bs\}$       
\\[2ex]

@@

% -------------------------------------
\paragraph{Acyclic and precedence-preserving} The first group of constraints ensures that the clustering is acyclic:
\begin{tabbing}
MM  \= MMMx \= M \= MMM \= M \= MMMM \= \kill
    \> ~~~~~~~ $x_{i,j}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N.\; x_{i,j}$ 
    \>             (with an edge from $i$ to $j$)            \\
    \> $-N.\;  x_{i,j}$  \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N.\; x_{i,j}$ 
    \>             (with no edge from $i$ to $j$)
\end{tabbing}
As per Megiddo~\cite{megiddo1998optimal} the form of these constraints is determined by whether there is an dependency between nodes $i$ and $j$. The $N$ value is set to the total number of nodes in the graph.

@@

% -----------------------------------------------------------------------------
\subsection{The Objective Function}
\label{s:ObjectiveFunction}

@@

\begin{tabbing}
MMMMM   \= MMMM \= M \= \kill
Minimise   \>     \> $\Sigma_{(i,j) \in E} W_{ij} x_{ij}$   \\
           \> \> \> (memory traffic and loop overhead)         \\
           \> $+$ \> $\Sigma_{i \in V} N c_i$  \\
           \> \> \> (removing intermediate arrays)         \\
Subject to \> \ldots                                \\
Where      \>                                       \\
           \> $W_{ij} = N^2$ \> $~|$ \> $(i,j) \in E $         \\
           \> \> \> (fusing $i$ and $j$ will reduce memory traffic)         \\
           \> $W_{ij} = N^2$ \> $~|$ \> $\exists k. (k,i) \in E \wedge (k,j) \in E $     \\
           \> \> \> ($i$ and $j$ share an input array)                                         \\
           \> $W_{ij} = 1$   \> $~|$ \> $@otherwise@$                                                  \\
           \> \> \> (the only benefit is loop overhead)                                        \\
           \\
           \> $N = |V|$
\end{tabbing}

@@

For the graph, we define a function $split(a,b)$ to check whether there exists a fusion-preventing path between two nodes, $a$ and $b$.

\begin{tabbing}
MMM      \= MM   \=  MMMMMMMMMMMM    \=  \kill
$split$     \> @::@ \> $name \times name \to \mathbb{B}$      \\
MM        \= M    \= \kill
$split(a,b)$ \\
    \>$=$  \>$\forall p \in path(a,b) \cup path(b,a).\ \fusionpreventing \not\in p$\\
\end{tabbing}

With $split$ defined, we can refine our formulation to only generate constraints between two nodes if there is a chance they may be fused together.
The entire formulation of the integer linear program follows.

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> $\Sigma_{(i,j) \in E} W_{ij} x_{ij} + \Sigma_{i \in V} N c_i$  \\
           \> (if $split(i,j)$)         \\
Subject to \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j) \not\in E \wedge (j,i) \not\in E$)            \\
\\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j,\fusible) \in E$)     \\
\\
           \>             \>       \> $\pi_i < \pi_j$ \>       \>            \\
           \> (if $(i,j,\fusionpreventing) \in E$)    \\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (if $split(i,j)$) \\
           \> $c_{i }$    \> $ = $ \> $ 1 $           \>       \>            \\
           \> (if $\neg(split(i,j))$) \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \emptyset$)  \\
           \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \{i',j'\} \wedge$ \\

           \> \> $ split(i,j) \wedge split(i',j) \wedge split(i,j') \wedge split(i',j') $) \\
MMMMM   \= MMMM \= M \= \kill
Where      \>                                       \\
           \> $W_{ij} = N^2$ \> $~|$ \> $(i,j) \in E $         \\
           \> \> \> (fusing $i$ and $j$ will reduce memory traffic)         \\
           \> $W_{ij} = N^2$ \> $~|$ \> $\exists k. (k,i) \in E \wedge (k,j) \in E $     \\
           \> \> \> ($i$ and $j$ share an input array)                                         \\
           \> $W_{ij} = 1$   \> $~|$ \> $@otherwise@$                                                  \\
           \> \> \> (the only benefit is loop overhead)                                        \\
           \\
           \> $N = |V|$
\end{tabbing}

>>>>

\begin{figure}
\begin{tabbing}
MMMMM       \= M  \= \kill
$nodes$     \> @:@ \> $\function \to V$              \\
$edges$     \> @:@ \> $\function \to E$              \\
$edge$      \> @:@ \> $\{bind\} \times bind \to E$ 
\\[1ex]
$nodes(bs)$ \> $= \{(name(b), \iiter_{\Gamma,C}(b)) | b \in bs\}$       
\\[2ex]

@@

% -------------------------------------
\paragraph{Acyclic and precedence-preserving} The first group of constraints ensures that the clustering is acyclic:
\begin{tabbing}
MM  \= MMMx \= M \= MMM \= M \= MMMM \= \kill
    \> ~~~~~~~ $x_{i,j}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N \cdot x_{i,j}$ 
    \>             (with an edge from $i$ to $j$)            \\
    \> $-N \cdot  x_{i,j}$  \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N \cdot x_{i,j}$ 
    \>             (with no edge from $i$ to $j$)
\end{tabbing}
As per Megiddo~\cite{megiddo1998optimal} the form of these constraints is determined by whether there is an dependency between nodes $i$ and $j$. The $N$ value is set to the total number of nodes in the graph.

@@

% -----------------------------------------------------------------------------
\subsection{Objective Function}
\label{s:ObjectiveFunction}

@@

\begin{tabbing}
MMMMM   \= MMMM \= M \= \kill
Minimise   \>     \> $\Sigma_{(i,j) \in E} W_{ij} \cdot x_{ij}$   \\
           \> \> \> (memory traffic and loop overhead)         \\
           \> $+$ \> $\Sigma_{i \in V} N \cdot c_i$  \\
           \> \> \> (removing intermediate arrays)         \\
Subject to \> \ldots                                \\
Where      \>                                       \\
           \> $W_{ij} = N^2$ \> $~|$ \> $(i,j) \in E $         \\
           \> \> \> (fusing $i$ and $j$ will reduce memory traffic)         \\
           \> $W_{ij} = N^2$ \> $~|$ \> $\exists k. (k,i) \in E \wedge (k,j) \in E $     \\
           \> \> \> ($i$ and $j$ share an input array)                                         \\
           \> $W_{ij} = 1$   \> $~|$ \> $@otherwise@$                                                  \\
           \> \> \> (the only benefit is loop overhead)                                        \\
           \\
           \> $N = |V|$
\end{tabbing}

@@

For the graph, we define a function $possible(a,b)$ to check whether there exists no fusion-preventing path between two nodes, $a$ and $b$.
If there is no fusion-preventing path, then fusion between the two nodes is possible.
We also define a function $possible'(a,b)$, to check whether the pair has parents that may also be fused.

\begin{tabbing}
MMMM \= M \=     \= MMMMMMMMMMMM    \=  \kill
$possible$ \> $:$     \> $name \times name \to \mathbb{B}$      \\
MM        \= M    \= \kill
$possible(a,b)$ \\
    \>$=$  \>$\forall p \in path(a,b) \cup path(b,a).\ \fusionpreventing \not\in p$\\
\\
MMMM \= M \=     \= MMMMMMMMMMMM    \=  \kill
$possible'$     \> $:$ \> $name \times name \to \mathbb{B}$      \\
MM        \= M    \= \kill
$possible'(a,b)$ \\
    \>$=$  \>$\exists A, B.\  parents(a,b) = \{A,B\} \wedge possible(a,b)$ \\
    \>$\wedge$ \> $possible(A,a) \wedge possible(B,b) \wedge possible(A,B)$
\end{tabbing}

With $possible$ and $possible'$ defined, we can refine our formulation to only generate constraints between two nodes if there is a chance they may be fused together.
The entire formulation of the integer linear program follows.

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> $\Sigma_{(i,j) \in E} W_{ij} \cdot x_{ij} + \Sigma_{i \in V} N \cdot c_i$  \\
           \> (if $possible(i,j)$)         \\
Subject to \\
           \> $-N \cdot x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N \cdot x_{ij}$ \\
           \> (if $possible(i,j) \wedge (i,j) \not\in E \wedge (j,i) \not\in E$)            \\
\\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N \cdot x_{ij}$ \\
           \> (if $possible(i,j) \wedge (i,j,\fusible) \in E$)     \\
\\
           \>             \>       \> $\pi_i < \pi_j$ \>       \>            \\
           \> (if $(i,j,\fusionpreventing) \in E$)    \\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (if $possible(i,j)$) \\
           \> $c_{i }$    \> $ = $ \> $ 1 $           \>       \>            \\
           \> (if $\neg possible(i,j)$) \\
           \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge possible'(i,j)$) \\

\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge \neg possible'(i,j)$) \\
MMMMM   \= MMMM \= M \= \kill
Where      \>                                       \\
           \> $W_{ij} = N^2$ \> $~|$ \> $(i,j) \in E $         \\
           \> \> \> (fusing $i$ and $j$ will reduce memory traffic)         \\
           \> $W_{ij} = N^2$ \> $~|$ \> $\exists k. (k,i) \in E \wedge (k,j) \in E $     \\
           \> \> \> ($i$ and $j$ share an input array)                                         \\
           \> $W_{ij} = 1$   \> $~|$ \> $@otherwise@$                                                  \\
           \> \> \> (the only benefit is loop overhead)                                        \\
           \\
           \> $N = |V|$
\end{tabbing}
